---
title: "Reproducible Research: Peer Assessment 1"
author: "vmintam"
date: "01/17/2015"
output: html_document
---


## Loading and preprocessing the data
```{r}
unzip('activity.zip')
activity <- read.csv("activity.csv")
```

## What is the mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

```{r histogram1}
totalSteps.day <- aggregate(steps ~ date, data=activity, FUN=sum)
barplot(totalSteps.day$steps, names.arg=totalSteps.day$date, xlab="date", ylab="steps", col = 'blue', main = "Histogram of Total Steps taken per day")
```

2. Calculate and report the **mean** and **median** total number of
steps taken per day

 - calculate
```{r}
step_mean <- mean(totalSteps.day$steps)
step_median <- median(totalSteps.day$steps)
```

 - report
 
  **mean** total number of steps taken per day is `r step_mean`
  
  **median** total number of steps taken per day is `r step_median`

## What is the average daily activity pattern?

1. Make a time series plot (i.e. `type = "l"`) of the 5-minute
interval (x-axis) and the average number of steps taken, averaged
across all days (y-axis)

```{r histogram2}
averageSteps_interval <- aggregate(steps ~ interval, data=activity, FUN=mean)
plot(averageSteps_interval, type="l", xlab='5-minute interval', ylab='average number of steps', col='red', main='Average number of Steps Taken at different 5 minute Intervals')
```

2. Which 5-minute interval, on average across all the days in the
dataset, contains the maximum number of steps?

```{r}
locate_interval <- averageSteps_interval$interval[which.max(averageSteps_interval$steps)]
```

- that's `r locate_interval`th interval.

## Inputing missing values

1. Calculate and report the total number of missing values in the
dataset (i.e. the total number of rows with `NA`s)

```{r}
total_missing <- sum(is.na(activity))
```
- total `r total_missing` values missing in dataset.

2. Devise a strategy for filling in all of the missing values in the
dataset. The strategy does not need to be sophisticated. For
example, you could use the mean/median for that day, or the mean
for that 5-minute interval, etc.

I will use the means for the 5-minute intervals as fillers for missing
values.
```{r}
interval_steps <- function(interval){
        averageSteps_interval[averageSteps_interval$interval == interval, ]$steps
}
```
3. Create a new dataset that is equal to the original dataset but with
the missing data filled in.

```{r}
activity_filled <- activity
for (i in 1:nrow(activity_filled)){
        if (is.na(activity_filled[i, ]$steps)){
                activity_filled[i, ]$steps <- interval_steps(activity_filled[i, ]$interval)
        }
}
```

4. Make a histogram of the total number of steps taken each day and
Calculate and report the **mean** and **median** total number of
steps taken per day. Do these values differ from the estimates from
the first part of the assignment? What is the impact of imputing
missing data on the estimates of the total daily number of steps?

```{r}
NewTotalSteps_Day <- aggregate(steps ~ date, data=activity_filled, FUN=sum)
barplot(NewTotalSteps_Day$steps, names.arg=NewTotalSteps_Day$date, xlab="date", ylab="steps", main="total number of steps taken each day", col = 'green')
mean(NewTotalSteps_Day$steps)
median(NewTotalSteps_Day$steps)
```

- **mean** do not diff from the first
- **median** only differ a little
- The impact of the missing data seems rather low, at least when
estimating the total number of steps per day.


## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels --
        "weekday" and "weekend" indicating whether a given date is a
weekday or weekend day.

```{r}
daytype <- function(date) {
        if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
                "weekend"
        } else {
                "weekday"
        }
}
activity_filled$daytype <- as.factor(sapply(activity_filled$date, daytype))
```

2. Make a panel plot containing a time series plot (i.e. `type = "l"`)
of the 5-minute interval (x-axis) and the average number of steps
taken, averaged across all weekday days or weekend days
(y-axis).

```{r}
par(mfrow=c(2,1))
for (type in c("weekend", "weekday")) {
        steps.type <- aggregate(steps ~ interval,
                                data=activity_filled,
                                subset=activity_filled$daytype==type,
                                FUN=mean)
        plot(steps.type, type="l", main=type, col = 'pink')
}
```